---
title: "Core Quants Week 15 classwork"
author: "YOUR NAME"
date: "26/03/2020"
output: html_document
---

# Some probabilities

Some questions to answer, firstly ...

## ... on a Polish's village "demographic mystery"

In Miejsce Odrzanskie, no boys have been born for almost a decase - during that time, 12 girls were born. This has been reported widely (https://www.nytimes.com/2019/08/06/world/europe/poland-village-girls-no-boys.html). But just how surprising is that?

- How likely is it that this will be the case in Miejsce Odrzanskie for the next 12 babies, assuming a 50% chance of each baby being a boy?
- How likely is it that this will be the case in at least one of Polands 2,479 gminas (local councils, a unit much larger than a village)?
- What else would we need to consider to estimate how surprising that news story should be?

```{r}


```

## ... on "feeling the future"

In 2011, Garyl J. Bem published the research article ""Feeling the Future: Experimental Evidence for Anomalous Retroactive Influences on Cognition and Affect" that provided evidence that people can perceive the future before it has happened. The paper included 9 studies, 8 of which yielded significant results. Assuming, for a moment, that extra-sensory perception exists, 
* how likely is that result, assuming 80% power? Calculate that manually with the approach shared in the module notes.

```{r}
#Calculate the probability of getting 8 significant and 1 non-significant result manually:


#You can alternatively calculate the probability for each possible outcome like this
dbinom(1:9, size = 9, prob = 0.8)

#Bem found an effect size of d = .22 across his experiments and typically used 150 participants or fewer. What power that does give us, assuming the use of a simple t-test? How does that power change the calculation above?
pwr::pwr.t.test(n =, d = )

```

## ... on shared birthdays

Imagine being in a room and talking about when people's birthday is.
- how many people need to be there for you to have a 50% chance to find someone who shares your birthdays (just day and month, ignoring year)?
- what about a 95% chance?

Note that the easiest approach here is to calculate the probability for a given number of people, and then adjust that number up or down to get the result you want. Also remember that it is often easier to calculate the probability that something that does not happen at all rather than that it happens at least once.


```{r}

```

Now, if we are not thinking about you specifically, but about any two people sharing a birthday, how does the calculation differ?
- how many people need to be in a room to have a 50% chance for at least two people to share a birthday?
- how about a 95% chance?

Here lots of people are going to compare their birthdays. As a hint, with 10 people, there are 10*9/2 pairs that can meet and compare birthdays (no need to learn in detail about combinations and permutation, but if you are so inclined: https://betterexplained.com/articles/easy-permutations-and-combinations/)

```{r}

```

## On positive Covid tests

Medical tests are evaluated by sensitivity (how many positive cases are correctly identified) and specificity (how many negative cases are correctly identified).

Lateral flow tests have a sensitivity of ~76.8% and a specificity of ~99.68% [Source: Oxford University](https://www.ox.ac.uk/news/2020-11-11-oxford-university-and-phe-confirm-lateral-flow-tests-show-high-specificity-and-are)

1 in 340 people in England had Covid-19 last week [Source: ONS](https://www.ons.gov.uk/peoplepopulationandcommunity/healthandsocialcare/conditionsanddiseases/bulletins/coronaviruscovid19infectionsurveypilot/19march2021)

- Given (only) that population rate, how should you interpret a positive / negative lateral flow test result?

- How would that be different in a situation like December (1 in 85 people) versus August (1 in 1,900 people)?


```{r}

```

# Choosing statistical tests

What statistical tests would you use to test the following? What key R functions would you use?
- whether the share of countries in income groups (i.e. low, middle, high) differs significantly between continents?
- whether the strength of the link between GDP and life expectancy differs between continents?
- whether changes in fertility affect life expectancy through a reduction in infant mortality?
- Whether we can assume that data is missing completely at random, so that casewise deletion does not introduce bias?